---
title: "lab04_rhymond.1"
author: "rhymond.1"
date: "2024-02-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = NULL,
  fig.width = 5,
  fig.height = 4,
  set.seed(3202)
)

```

**Problem 1**

**part a**

$L(\mu , \sigma^2 ; x_1, ... , x_n$ = $\prod_{i=1}^{n}f(x_i;\mu,\sigma^2)$ = $\prod_{i=1}^{n}\frac{1}{\sigma\sqrt{2\pi}}exp\{\frac{(x_i-\mu)^2}{2\sigma^2}\}$ = $(2\pi\sigma^2)^\frac{-n}{2}exp\{\frac{\sum_{i=1}^{n}(x_i-\mu)^2}{2\sigma^2}\}$ = $(2\pi)^\frac{-n}{2}exp\{\frac{\sum_{i=1}^{n}(x_i-\mu)^2}{2}\}$ (because $\sigma^2$ = 1)

**part b**

```{r}
L_Normal <- function(data, mu) {
  n <- length(data)
  L <- sqrt(2 * pi * 1 ^ 2) ^ (-n / 2) * exp(-.5 * sum((data - mu) ^ 2))
  return(L)
}
```

**part c**

```{r}
frogs <- c(4.57, 4.73, 2.24, 4.16, 4.53, 3.79, 2.48, 4.74, 3.18)
(xbar <- mean(frogs))
```

**part d**

```{r}
mu_domain <- seq(from = 0, to = 7, by = .05)
Likelihood_values = sapply(mu_domain, L_Normal, data = frogs)
plot(
  mu_domain,
  Likelihood_values,
  xlab = "mu",
  ylab = "L(mu|X,sigma)",
  main = "Normal Likelihood of Frog Weights",
  type = "l",
  lwd = 2
)
```

**part e**

```{r}
mu_hat = mu_domain[which.max(sapply(mu_domain, L_Normal, data = frogs))]
print(mu_hat)

Normal_MLE <- function(mu) {
  sapply(mu, L_Normal, data = frogs)
}
optimize(f = Normal_MLE,
         interval = c(0, 7),
         maximum = TRUE)
```

**part f**

Maximum is the maximum value in the graph. It is the mode value for mu. Objective is the value from the optimize function.

**part g**

```{r}
logL_Normal <- function(data, mu) {
  Log_L = log(L_Normal(data, mu))
  return(Log_L)
}

Log_Likelihood_values = sapply(mu_domain, logL_Normal, data = frogs)

plot(
  mu_domain,
  Log_Likelihood_values,
  xlab = "mu",
  ylab = "logL(mu|X,sigma)",
  main = "Normal Log-Likelihood of Frog Weights",
  type = "l",
  lwd = 2
)
```

**part h**

```{r}
logNormal_MLE <- function(mu) {
  sapply(mu, logL_Normal, data = frogs)
}
optimize(f = Normal_MLE,
         interval = c(0, 7),
         maximum = TRUE)
```

**part i**

Log is monotonically increasing, therefore the max of both the likelihood and log-likelihood would be in the same place.

**Problem 2**

**part a**

$\prod_{i=1}^{n}f(x;\lambda)$ = $\prod_{i=1}^{n}\lambda e^{-\lambda x}$ = $(\lambda^n)(e^{-\lambda n})(e^{\sum_{i=1}^{n}x})$

**part b**

```{r}
L_Exponential <- function(data, lambda) {
  n <- length(data)
  L <- (lambda^n)*(exp(-lambda*n))*(exp(sum(data)))
  return(L)
}
```

**part c**

```{r}
metal <- c(0.51, 0.02, 0.15, 0.46, 0.11, 0.04, 0.39,0.52, 0.2, 0.17, 0.01, 0.02, 0.32, 1.37)
(xbar <- mean(metal))
```

**part d**

```{r}
lambda_domain <- seq(from = 0, to = 10, by = .05)
Likelihood_values = sapply(lambda_domain, L_Exponential, data = metal)
plot(
  lambda_domain,
  Likelihood_values,
  xlab = "lambda",
  ylab = "L(lambda|X)",
  main = "Normal Likelihood of Metal Concentrations",
  type = "l",
  lwd = 2
)
```
It appears to be skewed right but its maximum seems to be around 1.

**part e**

```{r}
Normal_MLE <- function(lambda) {
  sapply(lambda, L_Exponential, data = metal)
}
optimize(f = Normal_MLE,
         interval = c(0, 10),
         maximum = TRUE)
```

**part f**

```{r}
logL_Exponential <- function(data, lambda) {
  Log_L = log(L_Exponential(data, lambda))
  return(Log_L)
}

Log_Likelihood_values = sapply(lambda_domain, logL_Exponential, data = metal)

plot(
  lambda_domain,
  Log_Likelihood_values,
  xlab = "mu",
  ylab = "logL(mu|X,sigma)",
  main = "Normal Log-Likelihood of Frog Weights",
  type = "l",
  lwd = 2
)
```
The max seems to be around 1 like in the non-log graph.
'

**Problem 3**

```{r}
height <- c("Safe", "Safe", "Safe", "Safe", "Safe", "Safe", "Safe", "Safe", "Safe", "Flood", "Safe", "Safe", "Safe", "Flood", "Safe", "Safe", "Flood", "Flood", "Safe", "Safe", "Flood", "Safe", "Safe", "Safe", "Safe", "Safe", "Safe", "Safe", "Flood", "Safe")
height_numeric <- ifelse(height == "Safe", 0, 1)

L_Poisson <- function(data, lambda) {
  n <- length(data)
  L <- ((lambda^sum(data))*(exp(-lambda*n)))/(prod(factorial(data)))
  return(L)
}

lambda_domain <- seq(from = 0, to = 1, by = .05)
Likelihood_values = sapply(lambda_domain, L_Poisson, data = height_numeric)
plot(
  lambda_domain,
  Likelihood_values,
  xlab = "lambda",
  ylab = "L(lambda|X)",
  main = "Normal Likelihood of Metal Concentrations",
  type = "l",
  lwd = 2
)

Normal_MLE <- function(lambda) {
  sapply(lambda, L_Poisson, data = height_numeric)
}
optimize(f = Normal_MLE,
         interval = c(0, 1),
         maximum = TRUE)

logL_Poisson <- function(data, lambda) {
  Log_L = log(L_Poisson(data, lambda))
  return(Log_L)
}

Log_Likelihood_values = sapply(lambda_domain, logL_Poisson, data = metal)

plot(
  lambda_domain,
  Log_Likelihood_values,
  xlab = "mu",
  ylab = "logL(mu|X,sigma)",
  main = "Normal Log-Likelihood of Frog Weights",
  type = "l",
  lwd = 2
)
```

