---
title: "lab08_rhymond.1"
author: "rhymond.1"
date: "2024-03-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = NULL,
  fig.width = 5,
  fig.height = 4,
  set.seed(3202)
)

```

**Problem 1**

**part a**

Review the $(1-\alpha)100\%$ confidence interval for mean difference $\mu_x-\mu_y$ of two populations $X$ and $Y$ with normal data and an known variances $\sigma_y^2$ and $\sigma_x^2$.

$$(\bar{x}-\bar{y})\pm z_{1-\alpha_{/2}}\sqrt{\frac{\sigma_x^2}{n_x}+\frac{\sigma_y^2}{n_y}}$$

**part b**

Let $\mu_x=10$, $\sigma_x=3$, $n_x=5$, $\mu_y=5$, $\sigma_y=5.5$, $n_y=8$ and $\alpha=0.05$. 

Generate a large number of samples from these two population and the confidence intervals for $\mu_x-\mu_y$. Compute the coverage rate of these
intervals.

```{r}
n_x = 5
mu_x = 10
sigma_x = 3
n_y = 8
mu_y = 5
sigma_y = 5.5
alpha = 0.05
CI_Diff=function(n_x, mu_x, sigma_x, n_y, mu_y, sigma_y, alpha){
  sampel_x = rnorm(n_x, mu_x, sigma_x)
  sampel_y = rnorm(n_y, mu_y, sigma_y)
  xbar = mean(sampel_x)
  ybar = mean(sampel_y)
  sd = sqrt(sigma_x ^ 2 / n_x + sigma_y ^ 2 / n_y)
  
  LCL = (xbar - ybar) - qnorm(1 - alpha / 2) * sd
  UCL = (xbar - ybar) + qnorm(1 - alpha / 2) * sd
  coverage = (mu_x - mu_y) >= LCL & (mu_x - mu_y) <= UCL
  return(coverage)
}
NMC=10000
CIs=replicate(NMC,CI_Diff(n_x = n_x, mu_x = mu_x, sigma_x = sigma_x, 
                          n_y = n_y, mu_y = mu_y, sigma_y = sigma_y, 
                          alpha = alpha))
mean(CIs)
```

**part c**

Review the $(1-\alpha)100\%$ confidence interval for $\mu_x-\mu_y$ with normal data and  **unknown variances but they are equal, $\sigma_y^2=\sigma_x^2=\sigma^2$ with $\sigma^2$ is the common variance**.

$$(\bar{x}-\bar{y})\pm t_{(1-\alpha_{/2}, n_x+n_y-2)}s_p\sqrt{\frac{1}{n_x}+\frac{1}{n_y}}$$
where $s_p=\sqrt{\frac{(n_x-1)s_x^2+(n_y-1)s_y^2}{n_x+n_y-2}}$ is the pooled standard deviation.

In this part, we’ll create confidence intervals for $\mu_x-\mu_y$, in the context where data are normally distributed with
the **population variance are unknown but they are equal**. 

Let$\mu_x=10$, $\sigma_x=3$, $n_x=5$, $\mu_y=5$, $\sigma_y=3$, $n_y=8$ and $\alpha=0.05$. 
Generate a large number of samples and
the confidence intervals for $\mu_x-\mu_y$ with $\sigma_x$ and $\sigma_y$ are assumed to be **unknown but equal**. 

**(i)**

Compute the coverage rate of these intervals. 

```{r}
n_x = 5
mu_x = 10
sigma_x = 3
n_y = 8
mu_y = 5
sigma_y = 3
alpha = 0.05
CI_Diff=function(n_x, mu_x, sigma_x, n_y, mu_y, sigma_y, alpha){
  sampel_x = rnorm(n_x, mu_x, sigma_x)
  sampel_y = rnorm(n_y, mu_y, sigma_y)
  xbar = mean(sampel_x)
  sd_x = sd(sampel_x)
  ybar = mean(sampel_y)
  sd_y = sd(sampel_y)
  sp_2 = ((n_x - 1) * sd_x ^ 2 + (n_y - 1) * sd_y ^ 2) / (n_x + n_y - 2)
  # sp_2 is pooled variance
  
  LCL = (xbar - ybar) - qt(1 - alpha / 2, n_x + n_y - 2) *
    sqrt(sp_2) * sqrt(1 / n_x + 1 / n_y)
  UCL = (xbar - ybar) + qt(1 - alpha / 2, n_x + n_y - 2) * sqrt(sp_2) *
    sqrt(1 / n_x + 1 / n_y)
  coverage = (mu_x - mu_y) >= LCL & (mu_x - mu_y) <= UCL
  return(coverage)
}
NMC=10000
CIs=replicate(NMC, CI_Diff(n_x = n_x, mu_x = mu_x, sigma_x = sigma_x,
                           n_y = n_y, mu_y = mu_y, sigma_y = sigma_y, 
                           alpha = alpha))
mean(CIs)
```

**Calculate the pooled variance when the population variances are unknown but equal**

**(ii)**

Repeat part (i), but this time use `t.test()` to automatically compute the intervals for $\mu_x-\mu_y$ instead of doing to by hand to compute the coverage rate of these
intervals. 

To get the the lower and upper confidence limits from t.test(), use t.test()\$conf.int[1] and t.test()\$conf.int[2] to extract the lower and upper confidence limits receptively. 

```{r}
n_x = 5
mu_x = 10
sigma_x = 3
n_y = 8
mu_y = 5
sigma_y = 3
alpha=0.05
CI_Diff=function(n_x, mu_x, sigma_x, n_y, mu_y, sigma_y, alpha) {
  sampel_x = rnorm(n_x, mu_x, sigma_x)
  sampel_y = rnorm(n_y, mu_y, sigma_y)
  t = t.test(sampel_x,
             sampel_y,
             var.equal = TRUE,
             cinf.level = 1 - alpha)
  LCL = t$conf.int[1]
  UCL = t$conf.int[2]
  coverage = (mu_x - mu_y) >= LCL & (mu_x - mu_y) <= UCL
  return(coverage)
}
NMC = 10000
CIs=replicate(NMC, CI_Diff(n_x = n_x, mu_x = mu_x, sigma_x = sigma_x,
                           n_y = n_y, mu_y = mu_y, sigma_y = sigma_y, 
                           alpha = alpha))
mean(CIs)
```

**Problem 2**

**part a**

**(i)**

\[df = \frac{(\frac{s_x^2}{n_x}+\frac{s_y^2}{n_y})^2}{\frac{1}{n_x-1}(\frac{s_x^2}{n_x})^2+\frac{1}{n_y-1}(\frac{s_y^2}{n_y})^2}\]

**(ii)**

```{r}
n_x = 5
mu_x = 10
sigma_x = 3
n_y = 8
mu_y = 5
sigma_y = 5.5
alpha=0.05
CI_Diff=function(n_x, mu_x, sigma_x, n_y, mu_y, sigma_y, alpha) {
  sample_x <- rnorm(n_x, mu_x, sigma_x)
  sample_y <- rnorm(n_y, mu_y, sigma_y)
  
  t_result <- t.test(sample_x, sample_y, var.equal = FALSE, conf.level = 1 - alpha)
  LCL <- t_result$conf.int[1]
  UCL <- t_result$conf.int[2]
  
  coverage <- (mu_x - mu_y) >= LCL & (mu_x - mu_y) <= UCL
  
  return(coverage)
}
NMC = 10000
CIs=replicate(NMC, CI_Diff(n_x = n_x, mu_x = mu_x, sigma_x = sigma_x,
                           n_y = n_y, mu_y = mu_y, sigma_y = sigma_y, 
                           alpha = alpha))
mean(CIs)
```



**Problem 3**

```{r}
n_x = 5
mu_x = 10
sigma_x = 3
n_y = 8
mu_y = 5
sigma_y = 5.5
alpha=0.05
CI_Diff=function(n_x, mu_x, sigma_x, n_y, mu_y, sigma_y, alpha) {
  sampel_x = rnorm(n_x, mu_x, sigma_x)
  sampel_y = rnorm(n_y, mu_y, sigma_y)
  t = t.test(sampel_x,
             sampel_y,
             var.equal = FALSE,
             cinf.level = 1 - alpha)
  LCL = t$conf.int[1]
  UCL = t$conf.int[2]
  coverage = (mu_x - mu_y) >= LCL & (mu_x - mu_y) <= UCL
  return(coverage)
}
NMC = 10000
CIs=replicate(NMC, CI_Diff(n_x = n_x, mu_x = mu_x, sigma_x = sigma_x,
                           n_y = n_y, mu_y = mu_y, sigma_y = sigma_y, 
                           alpha = alpha))
mean(CIs)
```