---
title: "lab06_rhymond.1"
author: "rhymond.1"
date: "2024-03-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = NULL,
  fig.width = 5,
  fig.height = 4,
  set.seed(3202)
)

```

**Problem 1**

**part a**

**(i)**

$X \sim N(\mu, \frac{\sigma^2}{n})$

**(ii)**

E(X) = $\mu$, SD(X) = $\frac{\sigma}{\sqrt{n}}$

**part b**

**(i)**
```{r}
set.seed(3202)
NMC = 10000
n = 36
mu = 20
sigma = 3
xbar_MC = function(n, mu, sigma) {
  x = rnorm(n, mean = mu, sd = sigma)
  xbar = mean(x)
}
hist(replicate(NMC, xbar_MC(n, mu, sigma)),
     main = "Sampling distribution of sample mean",
     xlab = "Sample mean")
abline(v = 20, col = "red") # centered at mean
```

**(ii)**

Done

**(iii)**
```{r}
#let's take 20.5 as a value for sample mean
interval = c(20.5 - 0.96, 20.5 + 0.96)
print(interval)
```

**(iv)**
```{r}
#Find the quantile using qnorm()
qnorm((1 - 0.9) / 2)
```

**part c**
```{r}
n = 9
mu = 40
sigma = 6
xbar_MC = function(n, mu, sigma) {
  x = rnorm(n, mean = mu, sd = sigma)
  z = (mean(x) - mu) / (sigma / sqrt(n))
}
hist(replicate(NMC, xbar_MC(n, mu, sigma)),
     main = "Sampling distribution",
     xlab = "(Sample mean-mean)/(sigma/sqrt(n))")
```

```{r}
(mean = mean(replicate(NMC, xbar_MC(n, mu, sigma))))
(variance = var(replicate(NMC, xbar_MC(n, mu, sigma))))
```



**Problem 2**

**part a**

For a sufficiently large n, if you take n samples of a distribution, the n sample means will form a normal distribution centered approximately at the true mean.

**part b**

It will be a normal distribution if n is sufficiently large as the problem states.

**part c**

```{r}
n = 40
lambda = 8
xbar_MC = function(n, lambda) {
  x = rchisq(n, lambda)
  xbar = mean(x)
}
hist(replicate(NMC, xbar_MC(n, lambda)),
     main = "Sampling distribution",
     xlab = "Sample Mean")
```

It appears aproximately normal

**part d**

```{r}
n = 4
lambda = 8
xbar_MC = function(n, lambda) {
  x = rchisq(n, lambda)
  xbar = mean(x)
}
hist(replicate(NMC, xbar_MC(n, lambda)),
     main = "Sampling distribution",
     xlab = "Sample Mean")
```

It does not appear normal as n is not large enough.

**part e**

Because the central limit theorem needs a sufficently large n, typically at least 30, to hold true.

**part f**

```{r}
n = 40
lambda = 8
xbar_MC = function(n, lambda) {
  x = rchisq(n, lambda)
  z = (mean(x) - lambda) / (4 / sqrt(n))
}
hist(replicate(NMC, xbar_MC(n, lambda)),
     main = "Sampling distribution",
     xlab = "(Sample mean-lambda)/(4/sqrt(n))")
```

Yes it appears approximately normal.

```{r}
(mean = mean(replicate(NMC, xbar_MC(n, lambda))))
(variance = var(replicate(NMC, xbar_MC(n, lambda))))
```

**Problem 3**

**part a**

\[S = \sqrt{\frac{1}{n-1}\sum_{i=1}^{n}(X_i-\bar{X})^2}\]

**part b**

```{r}
n = 4
mu = 40
sigma = 6
xbar_MC = function(n, mu, sigma) {
  x = rnorm(n, mean = mu, sd = sigma)
  s = sqrt((1/(n-1)) * sum((x - mean(x))^2))
  z = (mean(x) - mu) / (s / sqrt(n))
}
hist(replicate(NMC, xbar_MC(n, mu, sigma)),
     main = "Sampling distribution",
     xlab = "(Sample mean-mean)/(S/sqrt(n))")
```

Yes it appears that it is bell-shaped, unimoal, and with larger tails.

**part c**

```{r}
n = 20
mu = 40
sigma = 6
xbar_MC = function(n, mu, sigma) {
  x = rnorm(n, mean = mu, sd = sigma)
  s = sqrt((1/(n-1)) * sum((x - mean(x))^2))
  z = (mean(x) - mu) / (s / sqrt(n))
}
hist(replicate(NMC, xbar_MC(n, mu, sigma)),
     main = "Sampling distribution",
     xlab = "(Sample mean-mean)/(S/sqrt(n))")
```

This appears more normal.

```{r}
n = 30
mu = 40
sigma = 6
xbar_MC = function(n, mu, sigma) {
  x = rnorm(n, mean = mu, sd = sigma)
  s = sqrt((1/(n-1)) * sum((x - mean(x))^2))
  z = (mean(x) - mu) / (s / sqrt(n))
}
hist(replicate(NMC, xbar_MC(n, mu, sigma)),
     main = "Sampling distribution",
     xlab = "(Sample mean-mean)/(S/sqrt(n))")
```

This appears even more normal.